\vspace{1.0cm} \LARGE \bf Reassortment between influenza B lineages and the emergence of a co-adapted PB1-PB2-HA gene complex

### 2013-12-03 fluB.tex
A few text changes... much more to go.

Like all members of \textit{Orthomyxoviridae}, influenza B viruses have segmented genomes, which allow viruses co-infecting the same cell to exchange segments (a process known as reassortment). 
Influenza A viruses are widely considered to be a major threat to human health worldwide due to the ability to cause pandemics in humans via reassortment of seasonal human and non-human influenza A strains. 
The only known persistent source of influenza B viruses are humans (and occasionaly seals \cite{osterhaus2000,bodewes2013}), thus limiting the virus's pandemic potential. 
Both influenza A and B evolve antigenically through time in a process known as antigenic drift, in which mutations to the haemagglutinin (HA) protein allow viruses to escape existing human immunity and persist in the human population, leading to recurrent seasonal epidemics \cite{bedford2013}.

@@

Previously, through the use of time-stamped phylogenetic trees, the reassortment history of major clades of influenza B has been analyzed \cite{chen2008}, showing extensive and often complicated patterns of reassortment between all segments of influenza B viruses both between and within Vic and Yam lineages.
\tbc{This analysis was limited because...}
However, another approach exists that provides a significantly cleaner look at reassortment patterns in influenza B.
In this approach, membership to either the Victoria or Yamagata lineage in the tree of one segment is used as a discrete trait in trees of other segments.
Reassortment events which result in the replacement of one segment lineage by another show up as changes in the discrete trait along a branch and allow the inference of parameters related to the subtree descended from the node displaying a changed trait (Figure \ref{methodFig}).
Here, we use discrete traits to model inter-lineage reassortments between Victoria and Yamagata lineage segments to look at lineage persistence times following reassortment events.

We show that despite extensive reassortment, PB1, PB2 and HA segments of influenza B viruses still survive as distinct Victoria and Yamagata lineages, which appear to be co-adapted to the point where virions which do not contain PB1, PB2 or HA segments derived entirely from either the Vic or the Yam lineage are selected against and, to the best of our knowledge, have not been sampled since 1996.
In other segments (PA, NP, NA, MP and NS) a single lineage has been fixed in the influenza B population: Yam for PA, NP, NA and MP and Vic for NS.

>>>>

Like all members of \textit{Orthomyxoviridae}, influenza B viruses have segmented genomes, which allow viruses co-infecting the same cell to exchange segments (a process known as reassortment). 
Influenza A viruses are widely considered to be a major threat to human health worldwide due to the ability to cause pandemics in humans via reassortment of seasonal human and non-human influenza A strains. 
The only known persistent host of influenza B viruses are humans (and occasionaly seals \cite{osterhaus2000,bodewes2013}), thus limiting the virus's potential to exhibit pandemics through the acquisition of antigenic novelty from an animal reservoir. 
Both influenza A and B evolve antigenically through time in a process known as antigenic drift, in which mutations to the haemagglutinin (HA) protein allow viruses to escape existing human immunity and persist in the human population, leading to recurrent seasonal epidemics \cite{bedford2013}.

@@

Previously, through the use of time-stamped phylogenetic trees, the reassortment history of major clades of influenza B has been analyzed \cite{chen2008}, showing extensive and often complicated patterns of reassortment between all segments of influenza B viruses both between and within Vic and Yam lineages.
\tbc{This analysis was limited because...}
Here, we introduce another approach that reveals a evolutionarily intriguing pattern of reassortment in influenza B.
In this approach, membership to either the Victoria or Yamagata lineage in the tree of one segment is used to label the individual viruses in the tree of the other segments.
Reassortment events which result in the replacement of one segment lineage by another show up as changes in the discrete trait along a branch and allow the inference of parameters related to the subtree descended from the node displaying a changed trait (Figure \ref{methodFig}).
Here, we use discrete traits to model inter-lineage reassortments between Victoria and Yamagata lineage segments to look at lineage persistence times following reassortment events.

We show that despite extensive reassortment, PB1, PB2 and HA segments of influenza B viruses still survive as distinct Victoria (Vic) and Yamagata (Yam) lineages, which appear to be co-adapted to the point where virions which do not contain PB1, PB2 or HA segments derived entirely from either the Vic or the Yam lineage are selected against and, to the best of our knowledge, have not been sampled since 1996.
In other segments (PA, NP, NA, MP and NS) a single lineage has been fixed in the influenza B population: Yam for PA, NP, NA and MP and Vic for NS.
